---
title: "Assignment-3"
author: "Jun Yan"
output:
  html_document:
    toc: true
    toc_float: true
---

### Start

Load stringr package and the data

```{r}
library(stringr)
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert" 
```

Extract the names from the raw data

```{r}
name <- unlist(str_extract_all(raw.data, "[[:alpha:],. ]{2,}")) 
name
```

### Problem 3.a) 

Step 1. Remove any words end with a period. This will remove all titles and second names

```{r}
name1 <- str_replace_all(name, pattern = "\\w+\\. ", replacement = "")
name1
```

Step 2. Create a function that split a name with comma and flip it so it's first_name first

```{r}
correctName <- function(string)
{
  spliting <- str_split(string, pattern = ", ")
  lname <- spliting[[1]][1]
  fname <- spliting[[1]][2]
  return(str_c(fname, lname, sep = " "))
}
```

Step 3. Use a for-loop to go thru all the names applying the function.

```{r}
for (i in 1:length(name1))
{
  if (str_detect(name1[i], pattern = ", "))
  { 
    name1[i] <- correctName(name1[i])
  }
}
name1
```

### Problem 3.b)

```{r}
title <- str_detect(name, pattern = "Rev.|Dr.")
data.frame(name, title)
```

Alternatively, below code will catch other titles as well, detecting pattern of words with 2 to 3 letters ending with a period.

```{r}
str_detect(name, pattern = "\\w{2,3}\\. ")
```

### Problem 3.c)

```{r}
second_name <- str_detect(name, pattern = " \\w{1}\\. ")
data.frame(name, second_name)
```

### Problem 4.a)

Numbers 0 thru 9 match one or more times ending with dollar sign $

```{r}
example4a <- "90210$"
str_detect(example4a, "[0-9]+\\$")
```

### Problem 4.b)

A word that is 1 to 4 lower case English letters in length

```{r}
example4b <- "test"
str_detect(example4b, "\\b[a-z]{1,4}\\b")
```

### Problem 4.c)

A .txt file named with or without any characters.

```{r}
example4c <- c("testing1.txt", ".txt")
str_detect(example4c, ".*?\\.txt$")
```

### Problem 4.d)

Date of birth formatted mm/dd/yyyy. Althought the query doesn't have to be for DOB.

```{r}
example4d <- c("07/02/2013", "12/34/5678")
str_detect(example4d, "\\d{2}/\\d{2}/\\d{4}")
```

### Problem 4.e)

This is for HTML code search.
A string surrounded by <> brackets, followed by string of any length more than 1, followed by back referencing the same string but added a slash /

```{r}
example4e <- c("<bold> Any string </bold>", "<646718>3</646718>")
str_detect(example4e, "<(.+?)>.+?</\\1>")
```

### Problem 9

Here's the string where there's a message hidden:

```{r}
message <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"
```

I noticed there are four types of characters in this string:

1. lower case letters [a-z]
2. upper case letters [A-Z]
3. digits [0-9]
4. two punctuation [[.]!]

My first plan was to remove these types one by one from the string to see what's the result.

```{r}
func <- c("[0-9]", "[a-z]", "[A-Z]", "[[.]!]")
str_replace_all(message, func, replacement = "")
```

Didn't look good. 

My second try was to remove combination of two of the four types from the string, then combination of three types. And check to see what's the result.

Here I use `combn` function to generate combinations, for example:

```{r}
combn(func, 2)
```

Here, it generates a matrix, where each column is a combination of the types.

I then wrote a double for-loop and used `str_c` to combine the types, and removed the combined types from the string message.

```{r}
combo <- combn(func, 2)
combo_dim <- dim(combo)

# Following double loops splice the func together
for (j in 1:combo_dim[2])
{
  new_func <- combo[1,j]
  for (i in 2:combo_dim[1])
  {
    new_func <- str_c(new_func, "|", combo[i,j])
  }
  
  # Now removing the combined funcs
  new_message <- str_replace_all(message, new_func, replacement = "")
  print(c(new_func, new_message))
}
```

I found the message! After removing the two types "[0-9]|[a-z]", the message appears to be *"Congratulations you are a supernerd"*.

I continued on to remove combination of three types, using the same method, just to see what happens. 

```{r}
combo <- combn(func, 3)
combo_dim <- dim(combo)

# Following double loops splice the func together
for (j in 1:combo_dim[2])
{
  new_func <- combo[1,j]
  for (i in 2:combo_dim[1])
  {
    new_func <- str_c(new_func, "|", combo[i,j])
  }
  
  # Now removing the combined funcs
  new_message <- str_replace_all(message, new_func, replacement = "")
  print(c(new_func, new_message))
}
```

As you can see, after removing "[0-9]|[a-z]|[[.]!]", the message still can be seen. This is because the message was hidden as upper case in the strings.

