---
title: "Assignment-2"
author: "Jun Yan"
output: html_document
---

### What is required to run this assignment?

1. 3 csv files containing the data:
    i) names.csv
    ii) movies.csv
     iii) ratings.csv
2. MySQL Workbench server setup and running
3. R package RMySQL installed

### Loading Data

I begin by establishing connection.

```{r eval=TRUE}
library(RMySQL)
con <- dbConnect(MySQL(), user = "user007", password = "123456789", host = "DESKTOP-1GKOF0M", dbname = "movierate")
```

The `user`, `password`, `host`, and `dbname` will depend on how your system is set up.

Now let's see what the tables look like.

```{r eval=TRUE}
dbGetQuery(con, "SELECT * FROM movies")
dbGetQuery(con, "SELECT * FROM names")
dbGetQuery(con, "SELECT * FROM ratings")
```

Notice that "Star Wars: The Last Jedi" is not out yet, so no one rated it, and in the `names` table, there is a dummy person named "No One" who has not rated any movies. 

### Join Table

I want to join the tables together. Here I use two `LEFT JOIN` statement.

```{r eval=TRUE}
querry1 <- "SELECT movie, last_name, first_name, rating
            FROM movies 
            LEFT JOIN ratings on movies.movieid = ratings.movieid
            LEFT JOIN names on ratings.nameid = names.nameid
            ORDER BY movie, rating DESC"

joinTable1 <- dbGetQuery(con, querry1)
joinTable1
```

Notice that the Star Wars movie is listed with NA as value. And the dummy person No One is not on this table. This is because `LEFT JOIN` preserves only the left side table of the join statement. 

Let's try `RIGHT JOIN`. It should preserve the right side table of the join statement.

```{r eval=TRUE}
querry2 <- "SELECT movie, last_name, first_name, rating
            FROM movies 
            RIGHT JOIN ratings on movies.movieid = ratings.movieid
            RIGHT JOIN names on ratings.nameid = names.nameid
            ORDER BY movie, rating DESC"

joinTable2 <- dbGetQuery(con, querry2)
joinTable2
```

As expected, Star Wars movie is dropped from the table, but the dummpy person No One is listed eventhough that person has not rated any movies.

### Average Rating

Here I am interested in finding the average rating of each movie, first using SQL, then compare with the result using R code.

**Using SQL**

```{r eval=TRUE}
querry <- "SELECT movie, AVG(rating) AS avg_rating
            FROM movies 
            LEFT JOIN ratings on movies.movieid = ratings.movieid
            LEFT JOIN names on ratings.nameid = names.nameid
            GROUP BY movie
            ORDER BY avg_rating DESC"

avgTable <- dbGetQuery(con, querry)
avgTable
```

**Using R**

```{r eval=TRUE}
tble <- aggregate(joinTable1$rating, by = list(joinTable1$movie), mean)
names(tble) <- c("movie", "avg_rating")
tble[order(tble$avg_rating, decreasing = TRUE), ]
```

The results are the same. 

