---
title: "Assignment-1"
author: "Jun Yan"
output:
  html_document: default
  pdf_document: default
---

### Loading Data

First I load the data directly from the URL address into a data.frame object.

```{r eval=TRUE}
theURL <- "https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data"
mushroomData <- read.table(theURL, header=FALSE, sep=",")
```

Let's get a feel what it looks like.

```{r eval=TRUE}
dim(mushroomData)
head(mushroomData)
tail(mushroomData)
```

It has `r dim(mushroomData)[1]` cases just like the data dictionary indicates. The head and tail of the data.frame matches what it looks like when you go to the URL. 

Let's check the number of cases of edible vs poison.

```{r eval=TRUE}
numEdible <- dim(subset(mushroomData, V1 == "e"))[1]
numEdible
numPoison <- dim(subset(mushroomData, V1 == "p"))[1]
numPoison
```

This matches with the numbers given by the data dictionary. So the loading is successful.

### Manipulating Data

The four columns I am interested in this data set are, `cap-color`, `gill-color`, `ring-type`, and `habitat`. Including the `edible-poison` column, I extract them from the `mushroomData`. From the data dictionary, they are label as `V1`, `V4`, `V10`, `V20`, and `V23` in the data set.

```{r eval=TRUE}
mushroom <- mushroomData[, c("V1", "V4", "V10", "V20", "V23")]
```

The columns are changed using `names` function.

```{r eval=TRUE}
names(mushroom) <- c("edible_poison", "cap_color", "gill_color", "ring_type", "habitat")
names(mushroom)
```

Now I replace the abbreviations in the data, according to the data dictionary. But first I want to use `table` function to save the composition in the data. So I can use it to compare after renaming the factors. This way I can check if there's anything wrong with the renaming method.

To do this, I create a for loop to apple `table` function to each column and save it in a list object.

```{r eval=TRUE}
beforeRename <- list()
for (i in 1:dim(mushroom)[2])
  {
  beforeRename[[i]] <- table(mushroom[,i])
  }
```

The abbreviations are now replaced.

```{r eval=TRUE}
levels(mushroom$edible_poison) <- list('edible'='e', 'poison'='p')
levels(mushroom$cap_color) <- list('brown'='n','buff'='b','cinnamon'='c','gray'='g','green'='r','pink'='p','purple'='u','red'='e','white'='w','yellow'='y')
levels(mushroom$gill_color) <- list('black'='k','brown'='n','buff'='b','chocolate'='h','gray'='g','green'='r','orange'='o','pink'='p','purple'='u','red'='e','white'='w','yellow'='y')
levels(mushroom$ring_type) <- list('cobwebby'='c','evanescent'='e','flaring'='f','large'='l','none'='n','pendant'='p','sheathing'='s','zone'='z')
levels(mushroom$habitat) <- list('grasses'='g','leaves'='l','meadows'='m','paths'='p','urban'='u','waste'='w','woods'='d')
```
Let's check if the dimension stays same.

```{r eval=TRUE}
dim(mushroom)
```
Now, let's check if the renaming is done right. I create another list of tables to compare with the tables before renaming.

```{r eval=TRUE}
afterRename <- list()
for (i in 1:dim(mushroom)[2])
  {
  afterRename[[i]] <- table(mushroom[,i])
  }
```
```{r eval=TRUE}
beforeRename
afterRename
```

It can be seen that the composition did not change. So the renaming is successful.

Interestly, from this list, we see that the `ring_type` composition has 8 factors according to the data dictionary: cobwebby=c,evanescent=e,flaring=f,large=l,none=n,pendant=p,sheathing=s,zone=z.

But the actual data does not use all factors. It does not use 3 factors: cobwebby, sheathing, and zone. In another word, none of the mushrooms have these 3 ring types.

Let's confirm using the original data set.

```{r eval=TRUE}
levels(mushroomData$V20)
```

So yes, the data dictionary listed 8 factors, while the data uses 5 factors.

### Finish Product

Below I present the finish product:

```{r eval=TRUE}
mushroom
```