---
title: "Assignment-3"
author: "Jun Yan"
output:
  html_document:
    toc: true
    toc_float: true
---

### Start

Load stringr package and the data

```{r}
library(stringr)
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert" 
```

Extract the names from the raw data

```{r}
name <- unlist(str_extract_all(raw.data, "[[:alpha:],. ]{2,}")) 
name
```

### Problem 3.a) 

Step 1. Remove any words end with a period. This will remove all titles and second names

```{r}
name1 <- str_replace_all(name, pattern = "\\w+\\. ", replacement = "")
name1
```

Step 2. Create a function that split a name with comma and flip it so it's first_name first

```{r}
correctName <- function(string)
{
  spliting <- str_split(string, pattern = ", ")
  lname <- spliting[[1]][1]
  fname <- spliting[[1]][2]
  return(str_c(fname, lname, sep = " "))
}
```

Step 3. Use a for-loop to go thru all the names applying the function.

```{r}
for (i in 1:length(name1))
{
  if (str_detect(name1[i], pattern = ", "))
  { 
    name1[i] <- correctName(name1[i])
  }
}
name1
```

### Problem 3.b)

```{r}
title <- str_detect(name, pattern = "Rev.|Dr.")
data.frame(name, title)
```

Alternatively, below code will catch other titles as well, detecting pattern of words with 2 to 3 letters ending with a period.

```{r}
str_detect(name, pattern = "\\w{2,3}\\. ")
```

### Problem 3.c)

```{r}
second_name <- str_detect(name, pattern = " \\w{1}\\. ")
data.frame(name, second_name)
```

### Problem 4.a)

Numbers 0 thru 9 match one or more times ending with dollar sign $

```{r}
example4a <- "90210$"
str_detect(example4a, "[0-9]+\\$")
```

### Problem 4.b)

A word that is 1 to 4 lower case English letters in length

```{r}
example4b <- "test"
str_detect(example4b, "\\b[a-z]{1,4}\\b")
```

### Problem 4.c)

A .txt file named with or without any characters.

```{r}
example4c <- c("testing1.txt", ".txt")
str_detect(example4c, ".*?\\.txt$")
```

### Problem 4.d)

Date of birth formatted mm/dd/yyyy. Althought the query doesn't have to be for DOB.

```{r}
example4d <- c("07/02/2013", "12/34/5678")
str_detect(example4d, "\\d{2}/\\d{2}/\\d{4}")
```

### Problem 4.e)

This is for HTML code search.
A string surrounded by <> brackets, followed by string of any length more than 1, followed by back referencing the same string but added a slash /

```{r}
example4e <- c("<bold> Any string </bold>", "<646718>3</646718>")
str_detect(example4e, "<(.+?)>.+?</\\1>")
```






