---
title: "Project-1"
author: "Jun Yan"
output:
  html_document:
    toc: true
    toc_float: true
---

## Loading Data and Packages

First, begin by loading the data. Here, `readLines` function is used. This function turns the .txt file into a vector, where each line of the .txt is a element in the vector

```{r}
theURL <- "https://raw.githubusercontent.com/Tyllis/Data607/master/tournamentinfo.txt"
raw.data <- readLines(theURL)
```

Let's take a look the import result:

```{r}
head(raw.data)
tail(raw.data)
```

The import result can be verified by checking the length of the vector. The text file shows that there are 64 players, each player occupies 3 lines - the two lines containing each player's information and a line containing just dash lines "-". The text file also starts with 4 lines in the begining, containing two lines of just dash lines, and two lines containing the column names. So the total number of elements in the vector should be 64x3+4.

```{r}
length(raw.data) == 64*3+4
```

So the load is done correctly.

This project uses `stringr` package extensively.

```{r}
library(stringr)
```

## Vector Manipulation

A number of steps are needed to turn the vector into a data.frame object usable for analysis.

### Step 1. Remove the title elements

Here, the elements containing the titles are removed. we only need to look at the elements containing player and game information.

```{r}
vec1 <- raw.data[5:length(raw.data)]
head(vec1)
```

### Step 2. Remove dash line elements

In this step, the elements containing just dash lines "-" are removed.

```{r}
remove.value <- str_detect(vec1, pattern = "[-]{2,}")
vec2 <- vec1[!remove.value]
head(vec2)
```

Verifying that all dash lines are removed, there are one dash line for each player in the vector table1. So there are 64 total dash lines.

```{r}
length(vec2) == length(vec1) - 64
```

### Step 3. Combining elements

Now we have a vector that contains just player and game information. But each player's information is captured in two elements. We want to combine the two elements into one. 

This can be done by spliting the vector into two, one containing odd number reference and the other with even number reference. Then, the two vectors are spliced together using `str_c` function.

```{r}
split1 <- vec2[seq(1, length(vec2), 2)] 
split2 <- vec2[seq(2, length(vec2), 2)] 
vec3 <- str_c(split1, split2)
```

The result should be a 64 elements vector, where each element is a player.

```{r}
length(vec3)
```

Sampling around the vector just to get a feel what it looks like.

```{r}
vec3[sample(1:length(vec3), 10)]
```

### Step 4. Splitting elements 

Recognizing that each column in the text table is seperated by "|", we can use `str_split` to split the elements of the vector into a list, where each list element contains a vector, which in term contains a player's data. 

```{r}
lst4 <- str_split(vec3, pattern = "[|]")
lst4[1]
length(lst4) == 64
```

### Step 5. Trim spaces

In this step, the list is unpacked into vectors again, and spaces are trimmed off the string elements.

```{r}
vec5 <- str_trim(unlist(lst4))
```

Verifying the result by looking at the first and last player.

```{r}
vec5[1:21]
vec5[(length(vec5)-20):length(vec5)]
```

So each player's match data is now captured in 21 string elements. We should have 64x21 elements in the new vector.

```{r}
length(vec5) == 64*21
```

### Step 6. Extract the ratings

In this step, the rating data of each player is extracted using the `str_extract_all` function. Then, `str_detect` is used to flag the elements that contain the rating information, so it can be inserted back into the vector.

```{r}
ratings <- unlist(str_extract_all(vec5, pattern = "R:  ?[0-9]{3,4}"))
ratings_pos <- str_detect(vec5, pattern = "[0-9]{8} ")
vec6 <- vec5
vec6[ratings_pos] <- ratings
```

### Step 7. Remove Letters

In this step, the "W", "L", or "D" letters are removed from the string elements containing match information, so that only player numbers are left. The "R:" letters are also removed from the rating information.

```{r}
vec7 <- str_replace_all(vec6, pattern = "[WLD] |R: ", replacement = "")
```

Taking a look at the result:

```{r}
vec7[1:42]
vec7[(length(vec7)-20):length(vec7)]
length(vec7) == 64*21
```

## Creating `data.frame` Object

Now we can create the data.frame object. 

Recognizing that each type of data is seperated by 21 elements until the same type of data appears again in the vector, `seq` function can be used to reference all the same data in the vector. 

First, we create vectors containing the information we need:

```{r}
player_num <- as.numeric(vec7[seq(1, length(vec7), 21)])
player_name <- vec7[seq(2, length(vec7), 21)]
player_state <- vec7[seq(11, length(vec7), 21)]
points <- as.numeric(vec7[seq(3, length(vec7), 21)])
prerating <- as.numeric(vec7[seq(12, length(vec7), 21)])
round1 <- as.numeric(vec7[seq(4, length(vec7), 21)])
round2 <- as.numeric(vec7[seq(5, length(vec7), 21)])
round3 <- as.numeric(vec7[seq(6, length(vec7), 21)])
round4 <- as.numeric(vec7[seq(7, length(vec7), 21)])
round5 <- as.numeric(vec7[seq(8, length(vec7), 21)])
round6 <- as.numeric(vec7[seq(9, length(vec7), 21)])
round7 <- as.numeric(vec7[seq(10, length(vec7), 21)])
```

When `as.numeric` is used, if the string element contains non-numeric string, a "NA" is introduced to the data cell. Hence the warning. 

We now create the data.frame object:

```{r}
tour.data <- data.frame(player_num, player_name, player_state, points, prerating, round1, round2, round3, round4, round5, round6, round7)
str(tour.data)
head(tour.data)
tail(tour.data)
```

## Manipulating data.frame Object

The `round1` thru `round7` columns are referencing player numbers. We can use this information directly, by referencing them in the `prerating` column directly. We can fill the rounds columns with rating data instead.

```{r}
tour.data$round1 <- tour.data$prerating[tour.data$round1]
tour.data$round2 <- tour.data$prerating[tour.data$round2]
tour.data$round3 <- tour.data$prerating[tour.data$round3]
tour.data$round4 <- tour.data$prerating[tour.data$round4]
tour.data$round5 <- tour.data$prerating[tour.data$round5]
tour.data$round6 <- tour.data$prerating[tour.data$round6]
tour.data$round7 <- tour.data$prerating[tour.data$round7]

```

Now we can calculate the average match rating of opponents using `rowMeans` function.

```{r}
avgrating <- rowMeans(tour.data[,6:12], na.rm = TRUE)
round(avgrating)
tour.data$avgrating <- round(avgrating)
```

## Exporting .csv

Extract the columns needed for output.

```{r}
result <- tour.data[, c("player_name", "player_state", "points", "prerating", "avgrating")]
head(result)
tail(result)
```

Use `write.csv` to create .csv.

```{r}
write.csv(result, file = "proj-1-result.csv")
```

The .csv exported can be found here: https://raw.githubusercontent.com/Tyllis/Data607/master/proj-1-result.csv


