---
title: "DATA607 Week-5-Assignment"
author: "Jun Yan"
output: 
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 4
---

### Load Packages and Data

First, load the packages.

```{r}
library(tidyr)
library(dplyr)
```

The .csv data is loaded.

```{r}
theURL <- "https://raw.githubusercontent.com/Tyllis/Data607/master/airlinedata.csv"
aldata <- read.csv(theURL) 
```

Let's see what it looks like.

```{r}
aldata
```

### Transforming Data Set

The .csv is loaded into a data.frame object. There are a couple of initial issues:

- The csv data has a blank line in between the airlines. These blank line is imported as a line filled with "NA".
- The airline names are only showing in the "on time" row. They are omitted in the "delayed" row.

These two issues need to be dealt with first.

#### Removing "NA" line

The NA lines can be removed simply using the `filter` function.

```{r}
aldata <- filter(aldata, X != "")
```

#### Copy Airline Names

First we give the airline column a name.

```{r}
names(aldata)[1] <- c("Airline")
```

Now we can copy the airline names onto the "delayed" row.

```{r}
numrow <- dim(aldata)[1]
aldata[seq(2, numrow, 2), "Airline"] <- aldata[seq(1, numrow, 2), "Airline"]
aldata
```

The data.frame now looks more tidy than we started.

#### Cleaning Data

Upon inspection, this data has 4 variables:

- airline name
- city name
- number of on time flights
- number of delayed flights

The city names can be stacked using the `gather` function.

The number of on-time or delayed flights can be spread into two columns using `spread` function.

```{r}
aldata <- aldata %>% 
            gather(City, timliness, Los.Angeles:Seattle) %>% 
            spread(X, timliness)
aldata
```

It is now in its tidy form. Each column is a variable, and each row is a observation.

### Analyzing Data Set

I am interested to see which airline has more delays in these cities.

```{r}
aldata %>% 
  select(Airline, delayed) %>% 
  group_by(Airline) %>% 
  summarise(total_delays = sum(delayed))
```

So Am West has more delays. 

Let's look at the percentage or proportion of delays.

But first we need to change the column name "on time" to "on_time", so it doesn't throw an error when we use `dplyr` functions.

```{r}
names(aldata)[4] <- "on_time"
aldata %>% 
  select(Airline, delayed, on_time) %>% 
  group_by(Airline) %>% 
  summarise(delay_proportion = sum(delayed)/(sum(delayed+on_time)))
```

Percentage wise, Alaska Airline has more delay. 

Next, I am interested to see which city has the most delay by percentage.

```{r}
aldata %>% 
  group_by(City) %>% 
  summarise(delay_proportion = sum(delayed)/(sum(delayed+on_time)))
```

Therefore, San Francisco has 21.92% of its Alaska and Am West flights delayed, the highest among the cities.

Lastly, I would like to see, for each airline, which city has the most delay.

```{r}
aldata %>% 
  mutate(dp = delayed/(delayed+on_time)) %>% 
  group_by(Airline) %>% 
  filter(dp == max(dp))
```

Again, San Francisco tops the flight delay chart for both airlines.